Echoes of Economic Downturn: Investigating the Persistent Impact of the Great Recession on Birth Rates Among Young Americans

### 2024-02-15 outputs/paper/paper.qmd
update content

---
title: "Echoes of Economic Downturn: Investigating the Persistent Impact of the Great Recession on Birth Rates Among Young Americans"
author: 
  - Yuanyi (Leo) Liu
  - Qi Er (Emma) Teng
thanks: "Code and data are available at: https://github.com/leoyliu/Analyzing-the-Great-Recession-s-Impact-on-Young-Americans-Birth-Rates/tree/main. A replication of various aspects in this paper are available at: https://doi.org/10.48152/ssrp-srs6-t802"
date: today
date-format: long
abstract: "In this study, we explore the impact of the Great Recession on birth rates among young people in the United States, utilizing a comprehensive analysis of demographic and economic data. Our findings reveal a significant decline in birth rates within this group during and following the economic downturn, highlighting the intricate relationship between economic stability and reproductive decisions. This research contributes to our understanding of how macroeconomic factors can have profound effects on personal life choices, emphasizing the need for supportive policies targeting young individuals during economic crises."
format: html
toc: true
number-sections: true
bibliography: references.bib
---

@@

# install.packages("mapdata")
# install.packages("kableExtra")

@@

To better understand the data, the summary_table dataset was developed to offer a more structured and aggregated view of the birth rate data. By transforming and summarizing the data into a format that displays birth rate against years, this dataset simplifies the task of identifying and analyzing trends over time.

@@

rawdata |>
  select(year,brate_1519,brate_2024,brate_2529,brate_3034,brate_3539,brate_4044) |>
  slice(1:10) |>
  kable(
    caption = "Birth rate of all groups from 1980 to 1989",
    col.names = c("Year", "15 to 19", "20 to 24","25 to 29", "30 to 34", "35 to 39", "40 to 44"),
    digits = 1,
    booktabs = TRUE
  )
```

@tbl-display-raw-data uses R[@citekableExtra] presents the birth rates for various age groups in the United States from the year 1980 to 1989. Moreover, the birth rates are expressed per every 1,000 women in the population.To further understand this table, Year represent the years from the selected year range. The next six columns are different age groups for women, which are 15 to 19 years, 20 to 24 years, 25 to 29 years, 30 to 34 years, 35 to 39 years, and 40 to 44 years. Moreover, we are also concerned with the birth rate of related states in America.

@@

cleaned_data |>
  select(year, young, old) |>
  slice(1:10) |>
  kable(
    caption = "Birth rate of young and old age group from 1980 to 1989",
    col.names = c("Year", "Young", "Old"),
    digits = 1,
    booktabs = TRUE
  )
```

@tbl-display-raw-data uses R[@citekableExtra] presents the birth rates for the older and younger population from the year 1980 to 1989. This is a more concise table that represents the estimand we are trying to explore. Similarily to @tbl-display-raw-data, the birth rates are expressed by the mean birth rate of the selected age population. Moreover, the Year represent the years from the selected year range.

>>>>

---
title: "Echoes of Economic Downturn: Investigating the Persistent Impact of the Great Recession on Birth Rates Among Young Americans"
author: 
  - Yuanyi (Leo) Liu
  - Qi Er (Emma) Teng
thanks: "Code and data are available at: https://github.com/leoyliu/Analyzing-the-Great-Recession-Impact-on-Young-Americans-Birth-Rates. A replication of various aspects in this paper are available at: https://doi.org/10.48152/ssrp-srs6-t802"
date: today
date-format: long
abstract: "In this study, we explore the impact of the Great Recession on birth rates among young people in the United States, utilizing a comprehensive analysis of demographic and economic data. Our findings reveal a significant decline in birth rates within this group during and following the economic downturn, highlighting the intricate relationship between economic stability and reproductive decisions. This research contributes to our understanding of how macroeconomic factors can have profound effects on personal life choices, emphasizing the need for supportive policies targeting young individuals during economic crises."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

# install.packages("tidyverse")
# install.packages("dplyr")
# install.packages("here")
# install.packages("mapdata")
# install.packages("ggplot2")
# install.packages("haven")
# install.packages("kableExtra")

@@

To better understand the data, a summary table was developed to provide a detailed description of each variable, explaining its relevance and how it contributes to our understanding of the topic. Our focus is not only on the direct measures of birth rates but also on a range of demographic and economic variables that offer a comprehensive view of the factors influencing fertility decisions during the recessionary period.

@@

rawdata |>
  select(year,brate_1519,brate_2024,brate_2529,brate_3034,brate_3539,brate_4044) |>
  slice(1:10) |>
  kable(
    caption = "Birth rate of all groups from 1980 to 1989",
    col.names = c("Year", "15 to 19", "20 to 24","25 to 29", "30 to 34", "35 to 39", "40 to 44"),
    digits = 1,
    booktabs = TRUE
  )
```
@tbl-display-raw-data uses `kableExtra`[@citekableExtra] presents the birth rates for various age groups in the United States from the year 1980 to 1989. Moreover, the birth rates are expressed per every 1,000 women in the population.To further understand this table, Year represent the years from the selected year range. The next six columns are different age groups for women, which are 15 to 19 years, 20 to 24 years, 25 to 29 years, 30 to 34 years, 35 to 39 years, and 40 to 44 years. Moreover, we are also concerned with the birth rate of related states in America.

@@

cleaned_data |>
  select(year, young, old) |>
  slice(1:10) |>
  kable(
    caption = "Birth rate of young and old age group from 1980 to 1989",
    col.names = c("Year", "Young", "Old"),
    digits = 1,
    booktabs = TRUE
  )
```
@tbl-young-old-br uses `kableExtra`[@citekableExtra] presents the birth rates for the older and younger population from the year 1980 to 1989. This is a more concise table that represents the estimand we are trying to explore. Similarily to @tbl-display-raw-data, the birth rates are expressed by the mean birth rate of the selected age population. Moreover, the Year represent the years from the selected year range.
